---
title: "Arranging logger data from HoboLink"
author: "Jens Åström"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  NinaR::jensAnalysis:
    highlight: tango
    fig_caption: yes
    toc: yes
---


```{r, include = F}
#Some common packages, loading rmarkdown doesn't like the messages from tidyverse, so we don't include this in the document'
require(tidyverse)
require(DBI)
require(RPostgres)
require(ggplot2)
require(xtable)
require(NinaR)
```


```{r setup, include=FALSE}
#This is optional
#I choose the 'styler' package for tidying the code to preserve indentations
#I set the cutoff for code tidying to 60, but this doesn't currently work with styler.
#Set tidy = True to get the knitr default
#I want all figures as png and pdf in high quality in a subfolder called figure 

knitr::opts_chunk$set(echo = TRUE, 
                      tidy = "styler",
                      dev = c("png", "pdf"),
                      dpi = 600,
                      fig.path = "figure/",
                      tidy.opts = list(width.cutoff = 60)
                      )

options(xtable.comment = F, 
        xtable.include.rownames = F, 
        nina.logo.y.pos = 0.15)
palette(ninaPalette())
```



```{r, include = F, eval = F}
#This connects to the gisdatabase with a DBI connection named `con`.
#Use for example dbGetQuery(con, "SELECT * FROM ....") to query the database
postgreSQLConnect(host = "gisdata-db.nina.no", 
                 dbname = "gisdata", 
                 username = "postgjest", 
                 password = "gjestpost")
```

Intro
===========
The data exports for the temperature and humidity MX loggers from Hobo needs a bit of data wrangling before it can be used. The different data streams from each logger all get a separate column. Here we develop a script to turn this into a more usable long format.


Read in data
==========
I have a single export with many loggers, as a csv file. 

```{r}
inputFile <- "../rawData/All_MX_2020_2020_06_26_09_20_43_UTC_1.csv"

rawDat <- read_csv(inputFile,col_types = cols(.default = "c"))

dat <- rawDat %>%  
  select(-"Line#") %>% 
  mutate(date = as.POSIXct(Date, format = "%m/%d/%y %H:%M:%S")) %>% 
  mutate_if(is_character, as.double) %>% 
  select(-Date)

dat
```
That's quite the number of columns...



We have to pivot this data set to a longer format. We also get rid of the rows with no data.
```{r}
temp <- dat %>% 
  pivot_longer(cols = starts_with("Temperature"),
               names_to = "logger",
               values_to = "temperature") %>% 
  select(date,
         logger,
         temperature) %>% 
  filter(!is.na(temperature))

rh <- dat %>% 
  pivot_longer(cols = starts_with("RH"),
               names_to = "logger",
               values_to = "rh") %>% 
  select(date,
         logger,
         rh)%>% 
  filter(!is.na(rh))

dew  <- dat %>% 
  pivot_longer(cols = starts_with("Dew"),
               names_to = "logger",
               values_to = "dew") %>% 
  select(date,
         logger,
         dew) %>% 
  filter(!is.na(dew))

```

The data now looks like this
```{r}
temp
```

Time to strip the logger names and merge the tables

```{r}
temp <- temp %>% 
  mutate(logger = str_extract(logger,
                              "[^, ]+$"))

rh <- rh %>% 
  mutate(logger = str_extract(logger,
                              "[^, ]+$"))
dew <- dew %>% 
  mutate(logger = str_extract(logger,
                              "[^, ]+$"))

```

Check to see that the dates are the same for the datasets
```{r}
all(all(temp$date == rh$date),
all(rh$date == dew$date))
```

```{r}
combDat <- temp %>% 
  full_join(rh,
             by = c("date" = "date",
                    "logger" = "logger")) %>% 
  full_join(dew,
            by = c("date" = "date",
                    "logger" = "logger")) %>% 
  arrange(logger,
          date)
```


```{r}
combDat
```

Package this into a funtion
==========

```{r}
longerHobo <- function(inputFile){
  
  rawDat <- read_csv(inputFile,col_types = cols(.default = "c"))

  dat <- rawDat %>%  
    select(-"Line#") %>% 
    mutate(date = as.POSIXct(Date, format = "%m/%d/%y %H:%M:%S")) %>% 
    mutate_if(is_character, as.double) %>% 
    select(-Date)


    dat <- rawDat %>%  
  select(-"Line#") %>% 
  mutate(date = as.POSIXct(Date, format = "%m/%d/%y %H:%M:%S")) %>% 
  mutate_if(is_character, as.double) %>% 
  select(-Date)
 
   
  temp <- dat %>% 
    pivot_longer(cols = starts_with("Temperature"),
               names_to = "logger",
               values_to = "temperature") %>% 
    select(date,
         logger,
         temperature) %>% 
    filter(!is.na(temperature))

  rh <- dat %>% 
    pivot_longer(cols = starts_with("RH"),
                 names_to = "logger",
                 values_to = "rh") %>% 
    select(date,
           logger,
           rh)%>% 
    filter(!is.na(rh))
  
  dew  <- dat %>% 
    pivot_longer(cols = starts_with("Dew"),
                 names_to = "logger",
                 values_to = "dew") %>% 
    select(date,
           logger,
           dew) %>% 
    filter(!is.na(dew))
  
  
  temp <- temp %>% 
    mutate(logger = str_extract(logger,
                              "[^, ]+$"))
  rh <- rh %>% 
    mutate(logger = str_extract(logger,
                                "[^, ]+$"))
  dew <- dew %>% 
    mutate(logger = str_extract(logger,
                                "[^, ]+$"))
  
  if(!all(all(temp$date == rh$date),
  all(rh$date == dew$date))) stop("Tables datetimes doesn't match")
  
  combDat <- temp %>% 
  full_join(rh,
             by = c("date" = "date",
                    "logger" = "logger")) %>% 
  full_join(dew,
            by = c("date" = "date",
                    "logger" = "logger")) %>% 
  arrange(logger,
          date)
  
  return(combDat)
}
```

We can check that it produces the same results as the script.

```{r}
combDat2 <- longerHobo("../rawData/All_MX_2020_2020_06_26_09_20_43_UTC_1.csv")

all(combDat == combDat2)
```


Check the data out
=======

```{r}
ggplot(combDat) +
  geom_line(aes(x = date, y = temperature, color = logger)) +
  ggtitle("All the temperatures so far")

```

```{r}
oneLogger <- combDat %>% 
  filter(logger == "20835816") %>% 
  select(Date = date, 
         logger,
         Temperature = temperature,
         Relative_humidity = rh,
         Dew_point = dew) %>% 
  pivot_longer(-c(Date, logger),
               names_to = "Data_type",
               values_to = "Values")
  
ggplot(oneLogger) +
  geom_line(aes(x = Date, y = Values, color = Data_type)) +
  scale_color_nina() +
  ggtitle("All the data from one logger")
```



